Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2025-06-17 paper.qmd
Updated some texts

By default, when assessing the choice of $b_1$, the total number of bins is measured by the number of **non-empty** bins. This more accurately reflects the hexagon grid relative the RMSE than the full number of bins in the grid. It may be beneficial to remove low count bins also, such as when the data is clustered or stringy, or where the observed data is sparse. In order to decide if this is necessary, you would examine the distribution of bin counts, or the density which puts the counts on a standard scale. If there is something of a gap at low values, this would suggest a potential value to use as a cutoff. Alternatively, one could choose to remove based on a percentile, the bins with density in the lowest 5% of all bins, for example. @fig-param-two-curvy (c) illustrates the effect on the model representation of removing bins below different percentages. Generally, we would urge caution in removing low count bins.

The benchmark value for removing low-density hexagons ranges between $0$ and $1$, where $0$ corresponds to no bins being removed, and $1$ to 100% of bins being removed. When analyzing how these benchmark values influence model performance, it's essential to observe the change in RMSE as the benchmark value increases (@fig-param-two-curvy (c)). The RMSE shows a gradual decrease as the benchmark value goes from $1$ to $0$. Evaluating this rate of increase is important. If the increment is not considerable, the decision might lean towards retaining low-density hexagons.

@@

To understand how well the model fits the points whether it fits well, works better in some positions, or fails to match the overall pattern, it is important to link and brush the points with high model error in the \pD{} error plot, \gD{} layout, and the generated model overlay on the data in \pD{}.

>>>>

By default, when assessing the choice of $b_1$, the total number of bins is measured by the number of **non-empty** bins. This more accurately reflects the hexagon grid relative the RMSE than the full number of bins in the grid. It may also be beneficial to remove low count bins also, in the situation where data is clustered or stringy, where the observed data is sparse. In order to decide if this is necessary, you would examine the distribution of bin counts, or the density which puts the counts on a standard scale. If there is something of a gap at low values, this would suggest a potential value to use as a cutoff. Alternatively, one could choose to remove based on a percentile, the bins with density in the lowest $5\%$ of all bins, for example. @fig-param-two-curvy (c) illustrates the effect on the model representation of removing bins below different percentages. Generally, we would urge caution in removing low count bins.

The benchmark value for removing low-density hexagons ranges between $0$ and $1$. When analyzing how these benchmark values influence model performance, it's essential to observe the change in RMSE as the benchmark value increases (@fig-param-two-curvy (c)). The RMSE shows a gradual decrease as the benchmark value goes from $1$ to $0$. Evaluating this rate of increase is important. If the increment is not considerable, the decision might lean towards retaining low-density hexagons.

@@

To understand how well the model fits the points whether it fits well, works better in some positions, or fails to match the overall pattern. It is important to link and brush the points with high model error in the \pD{} error plot, \gD{} layout, and the generated model overlay on the data in \pD{}.
